---
title: "Health Insurence"
author: "Sudipta Das"
date: "2022-10-01"
output:
  word_document: default
  html_document: default
  pdf_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```



# Data Description

## Content

### Columns

     age: age of primary beneficiary

     sex: insurance contractor gender, female, male

     bmi: Body mass index, providing an understanding of body, weights that are relatively high or low relative to height,       objective index of body weight (kg / m ^ 2) using the ratio of height to weight, ideally 18.5 to 24.9

     children: Number of children covered by health insurance / Number of dependents

     smoker: Smoking

     region: the beneficiary's residential area in the US, northeast, southeast, southwest, northwest.

     charges: Individual medical costs billed by health insurance

```{r}
# Calling some libraries
library(ggplot2)
library(dplyr)
library(Hmisc)
library(cowplot)
library(WVPlots)
library(car)
library(caTools)
```


```{r}
# Calling my dataset

data=read.csv("C:/Users/Sudipta/Downloads/insurance.csv")
```

```{r}
head(data)
```
```{r}
describe(data)
```


### In our dataset there are some catagorical variables present. We want to convert them into numerical
```{r}
data$sex=factor(data$sex,levels = c("male","female"),labels = c(0,1))

```


```{r}
data$region=factor(data$region,levels=c("northeast","northwest","southeast","southwest"),labels = c(0,1,2,3))
```

```{r}
data$smoker=factor(data$smoker,levels = c("no","yes"),labels=c(0,1))
```


### Let's see our modified dataset

```{r}
head(data)
```

### EDA
```{r}
age<- ggplot(data, aes(age, charges)) +
  geom_jitter(color = "green", alpha = 0.7) +
    theme_dark()

bmi <- ggplot(data, aes(bmi, charges)) +
  geom_jitter(color = "green", alpha = 0.7) +
  theme_dark()

p <- plot_grid(age, bmi) 
title <- ggdraw() + draw_label("1. Plot of charges vs Age / BMI", fontface='bold')
plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))
```

```{r}
sex <- ggplot(data, aes(sex, charges)) +
  geom_jitter(aes(color = sex), alpha = 0.7) +
  theme_dark()

children <- ggplot(data, aes(charges,children)) +
  geom_jitter(aes(color = children), alpha = 0.7) +
  theme_light()

p <- plot_grid(sex, children) 
title <- ggdraw() + draw_label("2. Plot of  Charges vs Sex / Children ", fontface='bold')
plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))
```




### From plot of sex vs chaarges we get that charges of health insurence does not depende upon whether a person male or female.
```{r}
smoker<- ggplot(data, aes(smoker, charges)) +
  geom_jitter(aes(color = smoker), alpha = 0.7) +
  theme_dark()

region<- ggplot(data, aes(charges,region)) +
  geom_jitter(aes(color = region), alpha = 0.7) +
  theme_dark()

p <- plot_grid(smoker, region) 
title <- ggdraw() + draw_label("3. Plot of  Charges vs Smoker / Region", fontface='bold')
plot_grid(title, p, ncol=1, rel_heights=c(0.1, 1))
```



### As expected the charges of health insurence depends upon whether a person smoker or not. 
### It seems that changes in region has not that much effect on charges.



## Let's start Linear regression


### Splitting the dataset
```{r}
set.seed(1338)
n_train <- round(0.8* nrow(data))
train_indices <- sample(1:nrow(data), n_train)
data_train <- data[train_indices, ]
data_test <- data[-train_indices, ]
```


### Let's start model building
```{r}
## My first formula
formula_1 <- as.formula("charges ~ age + sex + bmi + children + smoker + region")
```


```{r}
## First model
model_1 <- lm(formula_1, data = data_train)
summary(model_1)

vif(model_1)
```
 
 
### So there is no multicolinearity in model_1

```{r}
## Let's predict data on test dataset using first model
prediction_1 <- predict(model_1, newdata = data_test)

#calculating the residuals
residuals_1 <- data_test$charges - prediction_1

#calculating Root Mean Squared Error
rmse_1 <- sqrt(mean(residuals_1^2))
rmse_1
```

```{r}
## My second formula
formula_2 <- as.formula("charges ~ age + bmi + children + smoker+region")

## Second model
model_2 <- lm(formula_2, data = data_train)
summary(model_2)

vif(model_2)
```


### So there is no multicolinearity in model_2
```{r}
## Let's predict data on test dataset using second model
prediction_2 <- predict(model_2, newdata = data_test)

## Calculating residuals
residuals_2 <- data_test$charges - prediction_2
rmse_2 <- sqrt(mean(residuals_2^2))
rmse_2
```

```{r}
## Here is my formula three
formula_3 <- as.formula("charges ~ age + bmi + children + smoker")

## Third model
model_3 <- lm(formula_3, data = data_train)
summary(model_3)

vif(model_3)
```


### Here also no multicolinearity in model_3
```{r}
## Let's predict the test dataset using third model
prediction_3 <- predict(model_3, newdata = data_test)

residuals_3 <- data_test$charges - prediction_3
rmse_3 <- sqrt(mean(residuals_3^2))
rmse_3
```


```{r}
r_sq_1 <- summary(model_1)$r.squared
r_sq_2 <- summary(model_2)$r.squared
r_sq_3 <- summary(model_3)$r.squared

print(paste0("R-squared for first model:", round(r_sq_1, 4)))
print(paste0("R-squared for second model:", round(r_sq_2, 4)))
print(paste0("R-squared for third model:", round(r_sq_3, 4)))

print(paste0("RMSE for first model: ", round(rmse_1, 2)))
print(paste0("RMSE for second model: ", round(rmse_2, 2)))
print(paste0("RMSE for third model: ", round(rmse_3, 2)))
```

###  As we can see, performance is quite similar between those models so we can drop the first model since rest two models  are little bit simpler.


### Now we have to decide whether we go with second or third model

### Performance the second model
```{r}
data_test$prediction1 <- predict(model_2, newdata = data_test)
ggplot(data_test, aes(x = prediction1, y = charges)) + 
  geom_point(color = "blue", alpha = 0.7) + 
  geom_abline(color = "green") +
  ggtitle("Prediction vs. Real values")+
  theme_dark()
```


```{r}
data_test$residuals1 <- data_test$charges - data_test$prediction1

ggplot(data = data_test, aes(x = prediction1, y = residuals1)) +
  geom_pointrange(aes(ymin = 0, ymax = residuals1), color = "blue", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = 3, color = "green") +
  ggtitle("Residuals vs. Linear model prediction")+
  theme_dark()
```


```{r}
ggplot(data_test, aes(x = residuals1)) + 
  geom_histogram(bins = 15, fill = "red") +
  ggtitle("Histogram of residuals")
```

```{r}
GainCurvePlot(data_test, "prediction1", "charges", "Model")
```


### Performance the third model
```{r}
data_test$prediction2 <- predict(model_3, newdata = data_test)
ggplot(data_test, aes(x = prediction2, y = charges)) + 
  geom_point(color = "blue", alpha = 0.7) + 
  geom_abline(color = "green") +
  ggtitle("Prediction vs. Real values")+
  theme_dark()
```






```{r}
data_test$residuals2 <- data_test$charges - data_test$prediction2

ggplot(data = data_test, aes(x = prediction2, y = residuals2)) +
  geom_pointrange(aes(ymin = 0, ymax = residuals2), color = "blue", alpha = 0.7) +
  geom_hline(yintercept = 0, linetype = 3, color = "green") +
  ggtitle("Residuals vs. Linear model prediction")+
  theme_dark()
```

```{r}
ggplot(data_test, aes(x = residuals2)) + 
geom_histogram(bins = 15, fill = "red") +
ggtitle("Histogram of residuals")


```

```{r}
GainCurvePlot(data_test, "prediction2", "charges", "Model")
```

### We can see the errors in both models are close to zero so both models predicts quite well.
### We will take the third model since it's simpler than the second model.

```{r}
## Calculating coefficients of third model
coef(model_3)
```

## So our predicted model is 
### Charges = -12258.5792 + 252.8285*(Age) + 331.3097*(BMI) + 479.27*(Children) + 23708.1619*(Smoker)
## With R^2 = 0.7616

# Conclusion:
    The aim of the project was to develop a model using multiple linear regression which can be used
    to predict health insurance charges. 
    
    Our predicted model is 
       Charges = -12258.5792 + 252.8285*Age + 331.3097*BMI + 479.27*(No of children) + 23708.1619*S
          where S=1 if the person is a smoker and S=0 if the person is not a smoker.












